\documentclass[draft]{NRELonepager}

\title{Concise Document}
\author{Andrew Clifton\thanks{Power Systems Engineering Center}}
\date{April 2017}

\addbibresource{bibliography.bib}

\begin{document}

\maketitle

\section{Using LaTeX to make documents that meet NREL's requirements}

A LaTeX class called \emph{NRELonepager.cls} has been written that implements NREL's formatting requirements in LaTeX, but is oriented towards shorter documents. This is not an official NREL RAAW.

This document is based on an extract from the \emph{NRELarticle.cls} demonstration document.

\subsection{NREL's LaTeX Environment}
NREL authors are strongly encouraged to use the NREL-hosted web-based latex environment to produce documents from LaTeX. This can be found at \href{latex.nrel.gov}{latex.nrel.gov} and is based on the open source interface used by \href{sharelatex.com}{sharelatex.com}. More information can be found at \href{latex.nrel.gov}{latex.nrel.gov}.

Advantages of \href{latex.nrel.gov}{latex.nrel.gov} include:
\begin{enumerate}
\item No need to maintain a local version of LaTeX
\item Everyone uses the same version of LaTeX
\item Editors and reviewers can work in a user-friendly environment
\item There is an always-on ``track changes'' feature
\item Secure hosting of documents within the NREL domain
\item Ability to download source documents for archiving
\end{enumerate}

\subsection{Starting new documents}
The NREL latex RAAWs are stored on NREL's Github account. To access and use those RAAWs:
\begin{enumerate}
\item Go to \href{https://github.com/NREL/latex_editing}{https://github.com/NREL/latex\_editing} and download the repository as a .zip file from the icon on the lower right hand side of the page.
\item Got to \href{latex.nrel.gov}{latex.nrel.gov} and start a new project by uploading the zip file. Modify the project properties (name, collaborators, etc) as required.
\item Navigate to the type of document that you want to use, e.g. article or report.
\item Modify \emph{main.tex} as required.
\end{enumerate}

\subsection{Functionality}
The \emph{NRELonepager.cls} style file has all of the functionality of the NREL article class and is based on the standard LaTeX article class. It uses reduced spacing for the title, section headings, and lists compared to the \emph{NRELarticle.cls}. Citations can be included if required.

This style file does not produce a tagged PDF document.

% bibliography
\label{sec:Bib}
\printbibliography

\end{document}
