\chapter{Preparing a .DOC or .DOCX file from LaTeX}\label{sec:latextodoc}
LaTeX users may find that they are required to convert their files into other formats for review or  editing. This section describes how LaTeX files can be converted into files that can be used with Microsoft Word.

\section{The conversion process} 
The NREL style has been designed to be converted into \emph{.doc} or \emph{.docx}. This can be achieved via one of two routes:
\begin{enumerate}
\item converting the latex file to rich text format (\emph{.rtf}) using \texttt{latex2rtf}, then importing the file into a word processor, or
\item converting the latex file to doc using pandoc.
\end{enumerate}

The performance of \texttt{Latex2rtf} and \texttt{pandoc} are compared in Table \ref{tab:conversioncomparison}.

\begin{table}[!h]
\centering
\caption{A comparison of latex2rtf and pandoc for document conversion.}\label{tab:conversioncomparison}
\begin{tabular}{lcc}
consideration & latex2rtf & pandoc \\\hline
requires changes to latex preamble & yes & no\\
output format & .rtf & .docx \\
supports cross references & yes & no\\
supports subfigures & partially & no\\
\end{tabular}
\end{table}

At this time, \texttt{latex2rtf} is the recommended tool for converting LaTeX files that have been made with the NREL class file, into word processor-readable files.

\section{Using latex2rtf to convert files}
The \texttt{latex2rtf} program reads LaTeX files and converts common LaTeX commands into their RTF equivalent. It is effectively another LaTeX interpreter that knows a limited subset of LaTeX. See the documentation at \href{http://sourceforge.net/projects/latex2rtf/}{http://sourceforge.net/projects/latex2rtf/} for details.

\subsection{Using latex2rtf}
To convert a document from LaTeX to RTF, follow these steps:
\begin{enumerate}
\item Install \texttt{latex2rtf}
\item Compile the document in LaTeX using the NREL class with the \texttt{book,report}, or \texttt{article} option, remembering to update the bibliography and cross references. The sequence of commands is:
\item Convert the document to RTF format using \texttt{latex2rtf}. The example .rtf file included with this document is created as follows:
\begin{verbatim}
latex2rtf -o latex2rtf_demo.rtf -f3 intro_to_NREL_latex
\end{verbatim}
\item Open the RTF file in Microsoft Word.
\item If the document contains tables of contents, tables of figures, tables of tables, or cross-references, select that text and update the fields.
\item Save the RTF file as a word-format document.
\end{enumerate}

\subsection{Using latex2rtf and LaTeX together}
Because \texttt{latex2rtf} only knows a subset of LaTeX, it is important to account for this when preparing a LaTeX document. The biggest problem is the lack of many packages, which is why authors are encouraged to use the NREL class file, which is known to work well with \texttt{latex2rtf}. Sometimes, though, it is important to be able to remove formatting for compatibility with \texttt{latex2rtf}, and so the preamble to this document includes a check to see if \texttt{latex2rtf} is being used:

\begin{verbatim}
\newif\iflatextortf
\iflatextortf
	\documentclass[12pt,letterpaper]{report}
	\input{NRELLatex2rtf.tex}
\else
	\documentclass[report]{nrel} 
\fi
\end{verbatim}

If \texttt{latex2rtf} is used, the boolean, \texttt{\textbackslash iflatextortf} will be TRUE and the commands will be interpreted as follows.
\begin{enumerate}
\item Set the document class to a generic LaTeX{} \emph{article}, \emph{report}, or \emph{book}. 
\item The file \emph{NRELLatex2rtf.tex} will be called, which maps most of the commands that are enabled in \emph{nrel.cls} to simpler versions that can be processed using \texttt{latex2rtf} (see Table \ref{Tab:Packages}).
\end{enumerate}

Authors that use packages other than those listed in Table \ref{Tab:Packages} may need to adjust the content of \emph{NRELLatex2rtf.tex} according to their needs. 

\subsection{Indexes}
Index entries will not be correctly converted to an \emph{.rtf} file. \emph{NRELLatex2rtf.tex} redefines the \verb+index+ command to do nothing when creating an \emph{.rtf} file. 

\subsection{What to do when the conversion to rich text format fails}
It is more than likely that the conversion to an \emph{.rtf} file will fail at some point. There are a few ways to deal with this:

\begin{description}
\item[Convert early and often.] Check that the document converts using \texttt{latex2rtf} every time a new environment is added.
\item[Try section-by-section.] Comment out the majority of the document and try to compile bit-by-bit. This will let you localize the error.
\item[Check new packages.] Please avoid using new packages. If a package has to be used, try the conversion immediately. If \texttt{latex2rtf} doesn't support the package, edit the file \emph{NRELLatex2rtf.tex} to redefine those commands to something that will convert appropriately. Put \emph{NRELLatex2rtf.tex} in the same directory as the LaTeX file to be converted.
\item[Avoid custom commands.] \texttt{latex2rtf} sometimes chokes on custom commands. A list of all recognized commands is available in the manual at \href{http://latex2rtf.sourceforge.net/latex2rtf.pdf}{http://latex2rtf.sourceforge.net/latex2rtf.pdf}. If custom commands are used, they may need to be redefined to work with the commands that \texttt{latex2rtf} does recognize. This can also be done in \emph{NRELLatex2rtf.tex}. You can check macros using the flag \verb+-d2+ when running \texttt{latex2rtf}.
\item[Use copy-paste.] Compile the whole document as a PDF, and save it somewhere. Then recompile using the reduced document that works with \texttt{latex2rtf}. Edit this in word and copy in the bits that killed the conversion.
\item[Talk to a communications rep.] If a document cannot be produced any other way than LaTeX with lots of packages, and \texttt{latex2rtf} just refuses to process it into a rich text file, discuss the process for having the PDF processed.
\end{description}

\section{Using pandoc}
Pandoc is a tool for converting documents between different formats. Pandoc is available at \href{http://johnmacfarlane.net/pandoc/}{http://johnmacfarlane.net/pandoc/}. Like \texttt{latex2rtf}, pandoc only supports a subset of latex commands and is not guaranteed to work for all applications.

Pandoc has several advantages over latex2rtf:
\begin{itemize}
\item Pandoc it supports conversion from latex to many other document formats (e.g. .html as well as .docx). However, there are other well-established tools to go directly from latex source code to other formats that may have better performance.
\item Pandoc does not require any changes to the latex source files.
\end{itemize}

To use pandoc, follow the installation instructions on the pandoc website, and then run pandoc from the command line. To convert this file, use the following command at the command line:

\begin{verbatim}
pandoc -o conversion_demo/pandoc_demo.docx intro_to_NREL_latex.tex --bibliography bibliography.bib --default-image-extension=.jpg
\end{verbatim}

